Catching fire: Undefined behaviour for equational reasoning of lazy languages with strictness annotations

### 2021-01-21 paper/paper.tex
blurb more
add fleshed out example 2

@@

Currently implementing the constructor simplification

@@

% =========================
\vspace{15em}

@@

\inference{
% foo
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{  \%xi = force(\%txi)} \\
\texttt{  \%zi = ... ; \%tzi = thunkify(\%zi)} \\
\texttt{  f(\%y1, ..., \%tzi, ..., \%yn)} \\
\texttt{\}}
}{
\texttt{func f\_strict\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = force(\%txi)} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
}
} % end footnotesize
\subcaption{outlining recursive call that is immediately forced}
\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize

\inference{
\texttt{func f(\%x1, ... \%wrapxi: @ADT, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
% \texttt{   \%zi = ...}
% \texttt{   \%wrapzi = construct(@Wrapper, \%zi)}
% \texttt{   f(\%y1, ..., \%wrapzi, ..., \%yn)}
\texttt{\}}
}{
\texttt{func f\_work\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_work\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
} % end inference
} % end footnotesize
\subcaption{outlining of recursion of a monovariant wrapper}
\end{minipage}
\end{tabular}

\vspace{15em}

@@

\vspace{15em}
\begin{tabular}{ll}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
}{
BAR
}
} % end footnotesize
\subcaption{Outline pattern matching branches on a function input}
% OutlineCaseOfFnInput
\end{minipage}
&

>>>>

Of course, much remains to be discussed: what about sum types? what about
non-tail recursion? Extensions to the same idea will handle the above problems,
while retaining the pleasing simplicity of this outlining/inlining paradigm.

@@

{\footnotesize
\begin{lstlisting}[language=c++, caption=Step 0: the initial program]
SimpleInt g(Thunk<SimpleInt> i) {
  SimpleInt icons = force(i);
  int ihash = casedefault(icons);
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    SimpleInt siprev = SimpleInt(prev);
    Thunk<SimpleInt> siprev_t = thunkify(siprev);
    SimpleInt g_prev_v = g(siprev_t);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 1: outline everything after \texttt{force}]
SimpleInt g2(SimpleInt i) {
  int ihash = casedefault(icons);
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    SimpleInt siprev = SimpleInt(prev);
    Thunk<SimpleInt> siprev_t = thunkify(siprev);
    SimpleInt g_prev_v = g(siprev_t);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}
SimpleInt g(Thunk<SimpleInt> i) {
  SimpleInt icons = force(i);
  g2(icons);
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 2: outline everything after \texttt{force}]
SimpleInt g2(SimpleInt i) {
  int ihash = casedefault(icons);
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    SimpleInt siprev = SimpleInt(prev);
    Thunk<SimpleInt> siprev_t = thunkify(siprev);
    SimpleInt g_prev_v = g(siprev_t);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}
SimpleInt g(Thunk<SimpleInt> i) {
  SimpleInt icons = force(i);
  g2(icons);
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 2: replace recursive call to \texttt{g}]
SimpleInt g2(SimpleInt i) {
  int ihash = casedefault(icons);
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    SimpleInt siprev = SimpleInt(prev);
    <@\textcolor{diffold}{ Thunk<SimpleInt> siprev\_t = thunkify(siprev)} @>;
    <@\textcolor{diffold}{ SimpleInt g\_prev\_v = g(siprev\_t);} @>
    <@\textcolor{diffnew}{ SimpleInt g\_prev\_v = g2(siprev);} @>
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}
SimpleInt g(Thunk<SimpleInt> i) {
  SimpleInt icons = force(i);
  g2(icons);
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 3: outline block after \texttt{casedefault}]
SimpleInt g3(int i) {
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    SimpleInt siprev = SimpleInt(prev);
    g_prev_v = g2(siprev);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}

SimpleInt g2(SimpleInt i) {
  int ihash = casedefault(icons);
  return g3(ihash);
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 4: replace \texttt{g2} with \texttt{g3}]
SimpleInt g3(int i) {
  if (ihash <= 0) {
    return SimpleInt(42);
  } else {
    int prev = ihash - 1;
    <@\textcolor{diffold}{SimpleInt siprev = SimpleInt(prev);} @>
    <@\textcolor{diffold}{g\_prev\_v = g2(siprev);}@>
    <@\textcolor{diffnew}{g\_prev\_v = g3(prev);}@>
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    SimpleInt ret = SimpleInt(rethash);
    return ret;
  }
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 5: float out \texttt{SimpleInt}]
SimpleInt g3(int ihash) {
  int out;
  if (ihash <= 0) {
    <@ \textcolor{diffold}{return SimpleInt(42)}; @>
    <@ \textcolor{diffnew}{out = 42;} @>
  } else {
    int prev = ihash - 1;
    SimpleInt g_prev_v = g3(prev);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    <@\textcolor{diffold}{SimpleInt ret = SimpleInt(rethash);} @>
    <@\textcolor{diffold}{return ret;}@>
    <@\textcolor{diffnew}{out = rethash;} @>
  }
  <@ \textcolor{diffnew}{return SimpleInt(out);} @>
}
\end{lstlisting}

\begin{lstlisting}[language=c++, caption=Step 5: taking stock]
SimpleInt g3(int ihash) {
  int out;
  if (ihash <= 0) {
    out = 42;
  } else {
    int prev = ihash - 1;
    SimpleInt g_prev_v = g3(siprev);
    int g_prev_v_hash = casedefault(g_prev_v);
    int rethash = g_prev_v_hash + 2;
    out = rethash;
  }
  return SimpleInt(out);
}
SimpleInt g2(SimpleInt i) {
  int ihash = casedefault(icons);
  return g3(ihash);
}

SimpleInt g(Thunk<SimpleInt> i) {
  SimpleInt icons = force(i);
  g2(icons);
}
\end{lstlisting}

@@

% =========================
\vspace{18em}

@@

\inference{
% foo
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{  \%xi = force(\%txi)} \\
\texttt{  \%zi = ... ; \%tzi = thunkify(\%zi)} \\
\texttt{  f(\%y1, ..., \%tzi, ..., \%yn)} \\
\texttt{\}}
}{
\texttt{func f\_strict\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = force(\%txi)} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
}
} % end footnotesize
\subcaption{outlining recursive call that is immediately forced}
\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
\texttt{func f(\%x1, ... \%wrapxi: @ADT, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
% \texttt{   \%zi = ...}
% \texttt{   \%wrapzi = construct(@Wrapper, \%zi)}
% \texttt{   f(\%y1, ..., \%wrapzi, ..., \%yn)}
\texttt{\}}
}{
\texttt{func f\_work\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_work\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
} % end inference
} % end footnotesize
\subcaption{outlining of recursion of a monovariant wrapper}
\end{minipage}
\end{tabular}

\vspace{20em}

@@

\vspace{18em}
\begin{tabular}{ll}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
}{
BAR
}
} % end footnotesize
\subcaption{Outline pattern matching branches on a function input}
% OutlineCaseOfFnInput
\end{minipage}
&
